%! Author = Kevin Lin
%! Date = 4/11/2025

% Preamble
\documentclass[11pt,a4paper,margin=1in]{article}

% Packages
\usepackage{amsmath}
\usepackage{enumerate}

\title{PSET 3}
\author{Kevin Lin}
\date{4/11/2025}

\begin{document}
\maketitle
\section{}
From class we know the hydrostatic balance equation:
\[
    \frac{dP}{dz} = -\rho g
\]
Let the pressure at the surface (also our lower boundary) be $P$. Because the
atmosphere is homogenous, we know $\rho$ is constant. Thus, reorganizing our
equation and integrating from the surface (0, $P_0$) to some height ($H$) (where
$P = 0$) gives us:
\begin{gather*}
    \int_{P_0}^{0} dP = -\rho g \int_{0}^{H}  dz\\
    0 - P_0 = -\rho g (H - 0)\\
    -P_0 = -\rho g H\\
    H = \frac{P_0}{\rho g}
\end{gather*}
At the surface, $P_0 = \rho R T_0$, where $T_0$ is our lower boundary temperature.
Thus:
\begin{gather*}
    H = \frac{\rho R T_0}{\rho g}\\
    H = \frac{R T_0}{g}
\end{gather*}
Therefore, $H$ must be a finite value as $R$ and $g$ are both constants, and 
$T_0$ is a constant determined at the lower boundary (surface), which is 
definitely not infinity.
\\\\
For a homogenous atmosphere with $T_0 = 298 \text{ K}$, the height is:
\begin{gather*}
    H = \frac{287 \text{ J kg}^{-1} \text{K}^{-1} \cdot 298 \text{K}}
        {9.8 \text{m s}^{-2}}\\
    H = 8727.14 \text{ m}
\end{gather*}

\section{}
\begin{enumerate}[A.]
\item
    Recall that the tropopause is the boundary between the troposphere and 
    stratosphere where the temperature changes from decreasing to increasing
    with height. We also know that pressure decreases with height. Combining
    these two facts, we can conclude that the troposphere is represented by the
    dotted line in the figure. From the figure, we see that the pressure goes
    from 200 mb to 1000 mb, which results in around 13km and that the 
    temperature changes from 300K to 200K. Thus, its lapse rate is:
    \begin{gather*}
        \Gamma = -\frac{dT}{dz} = -\frac{200\text{ K} - 300 \text{ K}}
            {13\text{ km}} = 7.7 \text{ K km}^{-1}
    \end{gather*}
    This is slightly lower than the dry adiabatic lapse rate. Note that this
    uses the a general estimate (described in TA OH), and not one of pi's and
    10's, which would result in a lapse rate of 10 K km$^{-1}$, equal to the
    dry adiabatic lapse rate.
\item
    The average temperature of the 700 mb to 300 mb layer is around 250K. Thus, 
    the pressure scale height is:
    \begin{gather*}
        H = \frac{RT_0}{g}\\
        H = \frac{287 \text{ J kg}^{-1} \text{K}^{-1} \cdot 250 \text{K}}
            {9.8 \text{m s}^{-2}}\\
        H \approx \frac{\pi \times 10^2 \times \pi \times 10^2}{10} \text{m}\\
        H \approx 10^4 \text{ m} 
    \end{gather*}
\end{enumerate}
\end{document}